Type-theoretical Semantics of \Pio\ via the Delay Monad \thanks{Supported by organization x.}

### 2019-04-24 mpc19/paper.tex
Hasegaws

\newcommand{\N}{\mathbb{N}}
\newcommand{\Pio}{$\mathsf{\Pi}^{\mathsf{o}}$} 
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Par}{\mathsf{Par}}
\newcommand{\comp}{\circ}
\newcommand{\U}{\mathcal{U}}
\newcommand{\copair}[2]{[#1,#2]}
\newcommand{\pair}[2]{\left< #1,#2 \right>}
\newcommand{\inl}{\mathsf{inl}}
\newcommand{\inr}{\mathsf{inr}}
\newcommand{\fst}{\mathsf{fst}}
\newcommand{\snd}{\mathsf{snd}}
\newcommand{\T}{\mathsf{T}}
\newcommand{\Inv}{\mathsf{Inv}}
\newcommand{\iter}{\mathsf{iter}}

@@

% Delay macros
\newcommand{\Delay}{\mathsf{Delay}\,}
\newcommand{\now}{\mathsf{now}}
\newcommand{\later}{\mathsf{later}}
\newcommand{\laterR}{\mathsf{laterR}}
\newcommand{\laterL}{\mathsf{laterL}}
\newcommand{\never}{\mathsf{never}}
\newcommand{\dn}{\downarrow}
\newcommand{\bind}{\mathsf{bind}}
\newcommand{\str}{\mathsf{str}}
\newcommand{\costr}{\mathsf{costr}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\Dapprox}{\mathbb{D}_{\approx}}
\newcommand{\copairD}[2]{[#1,#2]_{\D}}
\newcommand{\pairD}[2]{\left< #1,#2 \right>_{\D}}
\newcommand{\inlD}{\mathsf{inl}_{\D}}
\newcommand{\inrD}{\mathsf{inr}_{\D}}
\newcommand{\fstD}{\mathsf{fst}_{\D}}
\newcommand{\sndD}{\mathsf{snd}_{\D}}
\newcommand{\piso}{\mathsf{isPartialIso}}
\newcommand{\pisoalt}{\mathsf{isPartialIso}_2}
\newcommand{\iterD}{\mathsf{iter}_\D}
\newcommand{\unfolding}{\mathsf{unfolding}}
\newcommand{\refl}{\mathsf{refl}}

@@

@@

@@

>>>>

\newcommand{\N}{\mathbb{N}}
\newcommand{\Pio}{$\mathsf{\Pi}^{\mathsf{o}}$} 
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\Par}{\mathsf{Par}}
\newcommand{\comp}{\circ}
\newcommand{\U}{\mathcal{U}}
\newcommand{\copair}[2]{[#1,#2]}
\newcommand{\pair}[2]{\left< #1,#2 \right>}
\newcommand{\inl}{\mathsf{inl}}
\newcommand{\inr}{\mathsf{inr}}
\newcommand{\fst}{\mathsf{fst}}
\newcommand{\snd}{\mathsf{snd}}
\newcommand{\T}{\mathsf{T}}
\newcommand{\Inv}{\mathsf{Inv}}
\newcommand{\iter}{\mathsf{iter}}
\newcommand{\mapL}[1]{#1_{\mathsf{L}}}
\newcommand{\mapR}[1]{#1_{\mathsf{R}}}

@@

% Delay macros
\newcommand{\Delay}{\mathsf{Delay}\,}
\newcommand{\now}{\mathsf{now}}
\newcommand{\later}{\mathsf{later}}
\newcommand{\laterR}{\mathsf{laterR}}
\newcommand{\laterL}{\mathsf{laterL}}
\newcommand{\never}{\mathsf{never}}
\newcommand{\dn}{\downarrow}
\newcommand{\bind}{\mathsf{bind}}
\newcommand{\str}{\mathsf{str}}
\newcommand{\costr}{\mathsf{costr}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\Dapprox}{\mathbb{D}_{\approx}}
\newcommand{\copairD}[2]{[#1,#2]_{\D}}
\newcommand{\pairD}[2]{\left< #1,#2 \right>_{\D}}
\newcommand{\inlD}{\mathsf{inl}_{\D}}
\newcommand{\inrD}{\mathsf{inr}_{\D}}
\newcommand{\fstD}{\mathsf{fst}_{\D}}
\newcommand{\sndD}{\mathsf{snd}_{\D}}
\newcommand{\piso}{\mathsf{isPartialIso}}
\newcommand{\pisoalt}{\mathsf{isPartialIso}_2}
\newcommand{\iterD}{\mathsf{iter}_\D}
\newcommand{\unfolding}{\mathsf{unfolding}}
\newcommand{\refl}{\mathsf{refl}}
\newcommand{\traceD}{\mathsf{trace}_\D}
\newcommand{\traceHD}{\mathsf{traceH}_\D}

@@

The operation $\iterD$ is compatible with weak bisimilarity, which means
that $\iterD\,f_1 \approx \iterD\,f_2$ whenever $f_1 \approx f_2$.

@@

From the Elgot iteration operator it is possible to derive a trace
operator. First, given $f : A + B \to C$, we introduce
$\mapL f = f \comp \inl : A \to C$ and
$\mapR f = f \comp \inr : B \to C$, so that $f = [\mapL f , \mapR f]$.
The trace operator in $\Dapprox$ is defined as follows:
\begin{align*}
& \traceD : (A + B \to \Delay (A + C)) \to B \to \Delay C \\
& \traceD \,f = \copairD{\iterD \,\mapL f}{\now} \diamond \mapR f
\end{align*}
The operation $\traceD$ is compatible with weak bisimilarity.

It is well-known that a trace operator is obtainable from an iteration
operator. This is how Hasegawa's construction  \cite{Hasegwa97} derives it in the following way:
\begin{align*}
& \traceD : (A + B \to \Delay (A + C)) \to B \to \Delay C \\
& \traceD \,f = \copairD{\iterD \,\mapL f}{\now} \diamond \mapR f
\end{align*}

@@

%\ctikzfig{language}
